---
output:
  html_document: 
    toc: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

libs <- c('data.table', 'DT', 'tidyr')
lapply(libs, require, character.only = TRUE)

## Input data ----
results.path <-  file.path('Data', 'results')

census_yr <- fread(file.path('Data', 'herds_last_census.csv'))

## load results ----
files <- list.files(path = results.path)
results <- data.table(files = files[files %like% 'results'])
results[,names := stringr::str_extract(files, "(?<=results_).*(?=_202407)")]

for (i in 1:length(results$files)) { # for each file in the list
  fileName <- results$files[[i]] # save filename of element i
  dataName <- results$names[[i]] # save data name of element i
  tempData <- fread (file.path(results.path, fileName)) # read csv file
  assign (dataName, tempData, envir=.GlobalEnv)  # assign the results of file to the data named
  
}

dictionary <- data.table(cols = c("herd", 'last_census',
                                  "N.2024", "N.2024.uncertainty",
                         "lambda.2024", "lambda.2024.uncertainty",
                         "N.2027", "N.2027.uncertainty", 
                         "lambda.2027", "lambda.2027.uncertainty",
                         #"recovery5yrs", 
                         "time2recovery", 
                         "lambda.5yr", "lambda.5yr.uncertainty",
                         "cow.prod"),
                         labs = c("Subpopulation", 'Yr of last census',
                                  "# adult cows 2024", "% uncertainty in # cows 2024",
                                  "Lambda 2024", "% uncertainty lambda 2024",
                                  "# adult cows 2027", "% uncertainty in # cows 2027", 
                                  "Lambda 2027", "% uncertainty lambda 2027",
                                 # "% recovered in 5 yrs",
                                  "Yrs to recover pre-removal level", 
                                 "5-yr lambda", "% uncertainty 5-yr lambda",
                                 "Predicted cow growth"))

mk_table <- function(tab, title){
  tmp <- tab[,.(herd,          
         N.2027 = N.post_removal.median, N.2027.uncertainty = round((N.post_removal.median - N.post_removal.lower95)/N.post_removal.median*100),
         #lambda.2027 = round(exp(r.post_removal.median), 2), lambda.2027.uncertainty = round(abs(((exp(r.post_removal.median) - exp(r.post_removal.lower95))/exp(r.post_removal.median))*100)),
         lambda.5yr = round(exp(r.5yr.median), 2), lambda.5yr.uncertainty = round(abs(((exp(r.5yr.median) - exp(r.5yr.lower95))/exp(r.5yr.median))*100)),
         #recovery5yrs = round((N.2029.median/N.pre_removal.median)*100), 
         time2recovery.upper50 
         )]
  tmp[herd == 'A La Peche', herd:= 'À la Pêche']
  tmp[,time2recovery:=as.character(time2recovery.upper50)]
  tmp[is.na(time2recovery.upper50), time2recovery := '>100']
  tmp[time2recovery.upper50 == 2, time2recovery := '\u22642']
  tmp.order <- tmp[order(time2recovery.upper50, -lambda.5yr, lambda.5yr.uncertainty, -N.2027)]
  setnames(tmp.order, old = dictionary$cols, new = dictionary$labs, skip_absent = T)
  ft.tmp <- datatable(tmp.order[,-'time2recovery.upper50'], options = list(pageLength = 20))
  return(ft.tmp)
}
```

## Identifying source herds for Jasper National Park caribou captive breeding facility
* Authors:
    + [Julie W. Turner](https://wildlifescience.ca/team/julie-turner/)
    + [Tal Avgar](https://wildlifescience.ca/team/tal-avgar/)
    + [Robert Serrouya](https://wildlifescience.ca/team/dr-robert-serrouya/)

    
This repository contains the code for analyses [Biodiversity Pathways](https://biodiversitypathways.ca) did on behalf of Jasper National Park (JNP) to identify potential source herds of Southern Mountain Caribou to start their captive breeding program.

Below are tables from the report that can be sorted interactively to aid decision-making.

### Pre-removal status
Predicted 2024 herd statuses for herds based on the IPM and last known field censuses
```{r meta.tab, echo=FALSE}
meta.tab <- merge(noDD_remove0in2025_remove0in2026, census_yr, by = 'herd')
simp.meta.tab <- meta.tab[,.(herd, last_census = as.character(last_census),
                  N.2024 = N.pre_removal.median, N.2024.uncertainty = round((N.pre_removal.median - N.pre_removal.lower95)/N.pre_removal.median*100),
       lambda.2024 = round(exp(r.pre_removal.median), 2), 
       lambda.2024.uncertainty = round(abs(((exp(r.pre_removal.median) - exp(r.pre_removal.lower95))/exp(r.pre_removal.median))*100)),
       cow.prod = round(N.pre_removal.median*exp(r.pre_removal.median))-N.pre_removal.median)]
simp.meta.tab[herd == 'A La Peche', herd:= 'À la Pêche']
simp.meta.order <- simp.meta.tab[order(-cow.prod, -N.2024, -lambda.2024, lambda.2024.uncertainty, -last_census)]
setnames(simp.meta.order, old = dictionary$cols, new = dictionary$labs, skip_absent = T)

datatable(simp.meta.order[,-"Predicted cow growth"], options = list(pageLength = 20))

```


### Summary of years to recovery under each scenario
Predicted years to recovery combined across scenarios. Subpopulations are ranked by predicted cow growth in the 2024-2025 without removals, highlighting number of “surplus” cows.
```{r summary, echo=FALSE}
# remove duplicate 2024 vals
dups <- c("N.pre_removal.lower95", "N.pre_removal.median", "N.pre_removal.upper95", 
          "r.pre_removal.lower95", "r.pre_removal.median", "r.pre_removal.upper95" )
predResults <- rbind(noDD_remove3in2025_remove0in2026[,!(names(noDD_remove3in2025_remove0in2026) %in% dups), with = F][,removal := '3'], 
                    noDD_remove6in2025_remove0in2026[,!(names(noDD_remove6in2025_remove0in2026) %in% dups), with = F][,removal := '6'],
                     noDD_remove3in2025_remove3in2026[,!(names(noDD_remove3in2025_remove3in2026) %in% dups), with = F][,removal := '3/3'], 
                    noDD_remove12in2025_remove0in2026[,!(names(noDD_remove12in2025_remove0in2026) %in% dups), with = F][,removal := '12'],
                     noDD_remove6in2025_remove6in2026[,!(names(noDD_remove6in2025_remove6in2026) %in% dups), with = F][,removal := '6/6'], 
                    noDD_remove24in2025_remove0in2026[,!(names(noDD_remove24in2025_remove0in2026) %in% dups), with = F][,removal := '24'],
                     noDD_remove12in2025_remove12in2026[,!(names(noDD_remove12in2025_remove12in2026) %in% dups), with = F][,removal := '12/12'])
allResults <- merge(meta.tab[,.(herd, last_census = as.character(last_census),
                                N.2024 = N.pre_removal.median,
                                cow.prod = round(N.pre_removal.median*exp(r.pre_removal.median))-N.pre_removal.median)], 
                    predResults, 
                    by = 'herd')

yrs2recover <- allResults[,.(herd, last_census, N.2024, removal,
                             cow.prod, time2recovery.upper50)]
yrs2recover[,time2recovery := as.character(time2recovery.upper50)]
yrs2recover[is.na(time2recovery.upper50), time2recovery := '>100']
yrs2recover[time2recovery.upper50 == 2, time2recovery := '\u22642']
yrs2recover.wide <- dcast(yrs2recover, herd + last_census + N.2024 + cow.prod ~ removal, value.var = 'time2recovery')

yrs2recover.wide <- yrs2recover.wide[,.(herd, last_census, N.2024, cow.prod,
                                        `3`, `6`, `3/3`, `12`, `6/6`, `24`, `12/12`)]
yrs2recover.wide[herd == 'A La Peche', herd:= 'À la Pêche']

yrs2recover.order <- yrs2recover.wide[order(-cow.prod, -N.2024, -last_census)]
setnames(yrs2recover.order, old = dictionary$cols, new = dictionary$labs, skip_absent = T)
datatable(yrs2recover.order, options = list(pageLength = 20))
```


### Scenario: remove 3 cows
Predicted results of herd population viability when removing 3 cows in 2025 based on 1000 draws from the IPM posteriors
```{r rm3, echo=FALSE}
rm3 <- mk_table(noDD_remove3in2025_remove0in2026)
rm3

```


### Scenario: remove 6 cows
Predicted results of herd population viability when removing 6 cows in 2025 based on 1000 draws from the IPM posteriors
```{r rm6, echo=FALSE}
rm6 <- mk_table(noDD_remove6in2025_remove0in2026)
rm6

```


### Scenario: remove 6 cows staggered years
Predicted results of herd population viability when removing 3 cows in 2025 and 3 cows in 2026 based on 1000 draws from the IPM posteriors
```{r rm3n3, echo=FALSE}
rm3n3 <- mk_table(noDD_remove3in2025_remove3in2026)
rm3n3

```


### Scenario: remove 12 cows
Predicted results of herd population viability when removing 12 cows in 2025 based on 1000 draws from the IPM posteriors
```{r rm12, echo=FALSE}
rm12 <- mk_table(noDD_remove12in2025_remove0in2026)
rm12

```


### Scenario: remove 12 cows staggered years
Predicted results of herd population viability when removing 6 cows in 2025 and 6 cows in 2026 based on 1000 draws from the IPM posteriors
```{r rm6n6, echo=FALSE}
rm6n6 <- mk_table(noDD_remove6in2025_remove6in2026)
rm6n6

```


### Scenario: remove 24 cows
Predicted results of herd population viability when removing 24 cows in 2025 based on 1000 draws from the IPM posteriors
```{r rm24, echo=FALSE}
rm24 <- mk_table(noDD_remove24in2025_remove0in2026)
rm24

```


### Scenario: remove 6 cows staggered years
Predicted results of herd population viability when removing 3 cows in 2025 and 3 cows in 2026 based on 1000 draws from the IPM posteriors
```{r rm12n12, echo=FALSE}
rm12n12 <- mk_table(noDD_remove12in2025_remove12in2026)
rm12n12

```